---
title: "Lab09_Data606"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Grading the professor
Many college courses conclude by giving students the opportunity to evaluate the course and the instructor anonymously. However, the use of these student evaluations as an indicator of course quality and teaching effectiveness is often criticized because these measures may reflect the influence of non-teaching related characteristics, such as the physical appearance of the instructor. The article titled, “Beauty in the classroom: instructors’ pulchritude and putative pedagogical productivity” by Hamermesh and Parker found that instructors who are viewed to be better looking receive higher instructional ratings.

Here, you will analyze the data from this study in order to learn what goes into a positive professor evaluation.

# Getting Started

# Load packages

```{r }
library(tidyverse)
library(openintro)
library(GGally)
```

# The data

```{r }
glimpse(evals)
```

# Exercise 1: Is this an observational study or an experiment? The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations. Given the study design, is it possible to answer this question as it is phrased? If not, rephrase the question.

** Ans 1: The data are gathered from end of semester student evaluations for 463 courses taught by a sample of 94 professors from the University of Texas at Austin. The study is performed here beased on the students observation for their professors. As there is no experiment is involved here to get the data so obviously this ia an ovservational study. I think the beauty is not the direct parameter that can make differences in the course evaluations. Moreover, only six students rated on beauty parameter that is definitely not reflecting the whole class. So, the study question, " does beauty have an any effect on course evaluations?" should be the right question here.

```{r}
 hist(evals$score)
```
# Exercise 2: Describe the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not?

** Ans 2: The frequency distirbution of score is left sqewed. The center of the data will be described by the median of the data and which is greater than mean here. The distribution above depicts that the most of the scores are concetrated at the right side of the graph that means most of the students rated high scores to their professors which is a positive sign for professors evaluation. I always expect it as because high rates are the reflection of the good academic classes. 

# Exercise 3: Excluding score, select two other variables and describe their relationship with each other using an appropriate visualization.

** Ans 3: Let's see the relationship between the rank and age of the professors:
```{r }
plot(evals$rank, evals$age)
```
It is seen that those who are tenured are older than those who are in teaching and in tenure track. And those who are in tenure track are younger than others.

# Simple linear regression
The fundamental phenomenon suggested by the study is that better looking teachers are evaluated more favorably. Let’s create a scatterplot to see if this appears to be the case:

```{r }
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point()
```

Before you draw conclusions about the trend, compare the number of observations in the data frame with the approximate number of points on the scatterplot. Is anything awry?

# Exercise 4: Replot the scatterplot, but this time use geom_jitter as your layer. What was misleading about the initial scatterplot?

```{r }
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter()
```
** Ans 4: Well, the initial scatter plot drawn by geom_point did not add small amount of random variation to the location of each point and also did not overlap of points whereas scatter plot drawn by geom_jitter function did these things nicely. Thus it is reflecting a useful way of handling overplotting caused by discreteness in smaller datasets.

# Exercise 5: Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor score by average beauty rating. Write out the equation for the linear model and interpret the slope. Is average beauty score a statistically significant predictor? Does it appear to be a practically significant predictor?

** Ans5: Fitting a linear model m_bty to predict average professor score by average beauty rating:

```{r }
m_bty<-lm(evals$score~evals$bty_avg)
summary(m_bty)
```

So the equation for the linear model is: y = 0.06664 (bty_avg)+3.88034. The slope of the line is 0.06664 and it is  positive means the line's treand is upward. It means that the average score increases as the average beauty rating increases. Though the average beauty score appears to be a significant predictor, the R-squared value is 0.033 which is too low suggesting weak correlation between average beauty rating and professor average score. So, statiscally the predictor variable, average beauty rating does not have significant effect on professor average score. Practically it does not appear to be a significant predictor too. 

Add the line of the best fit model to the plot using the following:

```{r }

ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```
The blue line is the model. The shaded gray area around the line tells you about the variability you might expect in your predictions. To turn that off, use se = FALSE.

```{r }

ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)
```

# Exercise 6: Use residual plots to evaluate whether the conditions of least squares regression are reasonable. Provide plots and comments for each one (see the Simple Regression Lab for a reminder of how to make these).

** Ans 6: By using residual plots to evaluate whether the conditions of least squares regression are reasonable I will check: linearity between the two variables, nearly normal residuals and constant variability of residuals.

For lineartiy and constant varibility check I will plot the following residuals vs fitted values curve:

```{r }
ggplot(data = m_bty, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```
From the plot above, it is seen that the residuals are not distributed with a constant pattern around 0 
which does not indicate a linear trend. So, we can not assume a linear regression model for fitting the data set here.

To check the nearly normal residuals condition, we can look at the residuals histogram below:

```{r }
ggplot(data = m_bty, aes(x = .resid)) +
  geom_histogram(binwidth = 0.5) +
  xlab("Residuals")
```

It is seen that the histogram is not normally distributed (left-skewed). So, the residuals condition for being nearly normal is not met here.

Constant variability check: From the residuals vs. fitted plot above, it is seen that the spread of the residuals is not (roughly) equal at each level of the fitted values. So, we can say that the constant variance condition or assumption is not met here.

# Multiple linear regression
The data set contains several variables on the beauty score of the professor: individual ratings from each of the six students who were asked to score the physical appearance of the professors and the average of these six scores. Let’s take a look at the relationship between one of these scores and the average beauty score.

```{r }
ggplot(data = evals, aes(x = bty_f1lower, y = bty_avg)) +
  geom_point()
```
```{r}
evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```
As expected, the relationship is quite strong—after all, the average score is calculated using the individual scores.I can actually look at the relationships between all beauty variables (columns 13 through 19) using the following command:

```{r }
evals %>%
  select(contains("bty")) %>%
  ggpairs()
```

These variables are collinear (correlated), and adding more than one of these variables to the model would not add much value to the model. In this application and with these highly-correlated predictors, it is reasonable to use the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score after you’ve accounted for the professor’s gender, you can add the gender term into the model.

```{r }
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```
# Exercise 7: P-values and parameter estimates should only be trusted if the conditions for the regression are reasonable. Verify that the conditions for this model are reasonable using diagnostic plots.

**Ans 7: Linearity check:
```{r }
ggplot(data = m_bty_gen, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```
Residuals normality checK with histogram:

```{r }
ggplot(data = m_bty_gen, aes(x = .resid)) +
  geom_histogram(binwidth = 0.5) +
  xlab("Residuals")
```

or with normal probabity plot: 
```{r }
ggplot(data = m_bty_gen, aes(sample = .resid)) +
  stat_qq()
```
From the residuals plot above, it is seen that the linearity, nearly normal and constant variability conditions are met. So, it can be said that conditions for the regression model are reasonable

# Exercise 8: Is bty_avg still a significant predictor of score? Has the addition of gender to the model changed the parameter estimate for bty_avg?

**Ans 8: The bty_avg variable is still a significant predictor of score as because the p-value has decreased further with the addition of gender but it is still less than alpha (0.05).

Note that the estimate for gender is now called gendermale. we will see this name change whenever we introduce a categorical variable. The reason is that R recodes gender from having the values of male and female to being an indicator variable called gendermale that takes a value of 0 for female professors and a value of 1 for male professors. (Such variables are often referred to as “dummy” variables.)

As a result, for female professors, the parameter estimate is multiplied by zero, leaving the intercept and slope form familiar from simple regression.

scoreˆ=β^0+β^1×bty_avg+β^2×(0)=β^0+β^1×bty_avg


# Exercise 9 : What is the equation of the line corresponding to those with color pictures? (Hint: For those with color pictures, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which color picture tends to have the higher course evaluation score?

** Ans 9: I put in 1 for male parameter estimate and added 3.74734 and 0.17239 to get 3.91973. If two professors both receive the same beauty rating, the male professor will have the higher course evaluation score because 0.17239 is being added from the equation of the line based on the parameter estimate (0 for female and 1 for male).

The decision to call the indicator variable gendermale instead of genderfemale has no deeper meaning. R simply codes the category that comes first alphabetically as a 0. (we can change the reference level of a categorical variable, which is the level that is coded as a 0, using therelevel() function. Use ?relevel to learn more.)


# Exercise 10: Create a new model called m_bty_rank with gender removed and rank added in. How does R appear to handle categorical variables that have more than two levels? Note that the rank variable has three levels: teaching, tenure track, tenured.

** Ans 10: Each level of rank has separate line:
```{r }
m_bty_rank = lm(evals$score ~ evals$bty_avg + evals$rank)
summary(m_bty_rank)
```

It is seen from  the model summary above that there are two indicator variables: tenure track, and tenured. Here, teaching is a reference level to relatively interpret the indicator variables. This is why we don’t see “teaching level” in the regression output. The interpretation of the coefficients in multiple regression is slightly different from that of simple regression. The estimate for bty_avg reflects how much higher a group of professors is expected to score if they have a beauty rating that is one point higher while holding all other variables constant. In this case, that translates into considering only professors of the same rank with bty_avg scores that are one point apart.

# The search for the best model
We will start with a full model that predicts professor score based on rank, gender, ethnicity, language of the university where they got their degree, age, proportion of students that filled out evaluations, class size, course level, number of professors, number of credits, average beauty rating, outfit, and picture color.

# Exercise 11: Which variable would you expect to have the highest p-value in this model? Why? Hint: Think about which variable would you expect to not have any association with the professor score.

```{ r }
m_full <- lm(score ~ rank + gender + ethnicity + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```


** Ans 11: I am expecting that the cls_profs would have the highest p-value. This is because the number of teaching sections in a course is irrelevant to how teachers are evaluated. 

# Exercise 12: Check your suspicions from the previous exercise. Include the model output in your response.

**Ans 12: It is clearly seen from the model above that the p-value of ‘cls_profs’ is 0.77806 that is the highest value in the group of values.

# Exercise 13: Interpret the coefficient associated with the ethnicity variable.

** Ans 13: This value depicts that the non-minority professors score is 0.1234929 points higher than the minority professors.

# Exercise 14 : Drop the variable with the highest p-value and re-fit the model. Did the coefficients and significance of the other explanatory variables change? (One of the things that makes multiple regression interesting is that coefficient estimates depend on the other variables that are included in the model.) If not, what does this say about whether or not the dropped variable was collinear with the other explanatory variables?

```{r }
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

** Ans 14: It is seen that dropping the highest p-value, there is a small difference within the coefficients. This implies that the variables carry a greater significance.

# Exercise 15: Using backward-selection and p-value as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on.

** Ans 15: 
```{r }
m_backward <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_color, data = evals)
summary(m_backward)
```

score=3.771922+(ethnicity×0.167872)+(gender×0.207112)+(language×−0.206178)+(age×−0.006046)+(clsperceval×0.004656)+(clscreditsone×0.505306)+(btyavg×0.051069)+(piccolor×−0.190579)=3.91973+0.07416×bty_avg

# Exercise 16: Verify that the conditions for this model are reasonable using diagnostic plots.

** Ans 16: From the residuals plots below, it seems conditions are reasonable for this model.

```{r }
ggplot(data = m_backward, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```
```{r }
ggplot(data = m_backward, aes(x = .resid)) +
  geom_histogram(binwidth = 0.5) +
  xlab("Residuals")
```

```{r }
ggplot(data = m_backward, aes(sample = .resid)) +
  stat_qq()
```

# Exercise 17: The original paper describes how these data were gathered by taking a sample of professors from the University of Texas at Austin and including all courses that they have taught. Considering that each row represents a course, could this new information have an impact on any of the conditions of linear regression?

** Ans 17: If each row represents a course, there could be overlapping students skewing the data with their preferences.

# Exercise 18: Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.

** Ans 18: Based on my model the professors associated with a high evaluation score would young, non-minority, male, speaks English,higher amount of credits being taught, high number of evaluations, percieved as beautiful, and picture is colored.

# Exercise 19: Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?

** Ans 19: I would not be comfortable applying this conclusion generally because the adjusted R-squared value is very small (15.76%). Therefore, it is insignificant and thus reflecting a weak correlation. Moreover, it is not practical and also unfair to judge a professor by his/her appearance.


